---
title: "Initial Korean covid 19"
author: "Adam Ayala"
date: "10/3/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The majority of people who were alive female, at about 55%, while the rest were males, However, out of the population that died, nearly 65% were male while 45% were female. This means that there was a significant difference between the account if males that died and the amount of females that died, males being the majority and as nearly doubled for the amount of dead females. 

  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Females were shown to have 56 alive compared to males having 44 alive. Looking at the percentage we can see 1.4% for females and 3% for males meaning that there is a 2:1 ratio of males dying over females.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Null results expected that the percentage of the dead males and females compared to alive males and females would be equal. However, the observed counts showed that men nearly doubled the percentage of the female dead population. 


## Conclusions

The graph shows that a majority of males would die compared to females. To prove this checking the numbers we would see the percent doubled for males compared to females. Seeing if this was true with our null test we would see that the men dead group nearly doubled the percentage of the female dead group.